---
title: "Test 3D plots"
output: html_notebook
---
16.12.2021
Juan G. Diaz Ochoa

Test of 3D ploting ussing [this package](http://www.sthda.com/english/wiki/a-complete-guide-to-3d-visualization-device-system-in-r-r-software-and-data-visualization)

```{r}
library("rgl")
```


Download data
```{r}
data(iris)
head(iris)
```
Coordinates
```{r}
x <- sep.l <- iris$Sepal.Length
y <- pet.l <- iris$Petal.Length
z <- sep.w <- iris$Sepal.Width
```

scatter plot
```{r}
rgl.open() # Open a new RGL device
rgl.points(x, y, z, color ="lightgray") # Scatter plot
```

Same scatterplot with different shapes

```{r}
rgl.open()# Open a new RGL device
rgl.bg(color = "white") # Setup the background color
rgl.spheres(x, y, z, r = 0.2, color = "grey") 
```
scatter plot in a box. To this end a new device must be initialized

```{r}
#' @param new.device a logical value. If TRUE, creates a new device
#' @param bg the background color of the device
#' @param width the width of the device
rgl_init <- function(new.device = FALSE, bg = "white", width = 640) { 
  if( new.device | rgl.cur() == 0 ) {
    rgl.open()
    par3d(windowRect = 50 + c( 0, 0, width, width ) )
    rgl.bg(color = bg )
  }
  rgl.clear(type = c("shapes", "bboxdeco"))
  rgl.viewpoint(theta = 15, phi = 20, zoom = 0.7)
}
```

And then the data can be represented


```{r}
rgl_init()
rgl.spheres(x, y, z, r = 0.2, color = "yellow", xlabel = "x")  # Scatter plot
rgl.bbox(color = "#333377") # Add bounding box decoration
```

Alternative representation

```{r}
rgl_init()
plot3d(x, y, z, col="blue", type ="p")
```

```{r}
rgl_init()
plot3d(x, y, z, col="blue", box = FALSE,
       type ="s", radius = 0.15)
```

Test Results from persistent topology

```{r}
Phy_S_Diag_Mhealt_x <- read.csv("D:/Projects/2022_MANUSCRITOS/2020_SUMMA_BIOL/Notebooks_Programs/Results/Phy_S_Diag_Mhealt_x.csv")
Phy_S_Diag_Mhealt_y <- read.csv("D:/Projects/2022_MANUSCRITOS/2020_SUMMA_BIOL/Notebooks_Programs/Results/Phy_S_Diag_Mhealt_y.csv")
Phy_S_Diag_Mhealt_z <- read.csv("D:/Projects/2022_MANUSCRITOS/2020_SUMMA_BIOL/Notebooks_Programs/Results/Phy_S_Diag_Mhealt_z.csv")
```
Coordinates
```{r}
x <- Phy_S_Diag_Mhealt_x$Geom_Phi
y <- Phy_S_Diag_Mhealt_y$Geom_Phi
z <- Phy_S_Diag_Mhealt_z$Geom_Phi
```

3D representation
```{r}
rgl_init()
rgl.spheres(x, y, z, r = 0.09, color = "yellow", xlabel = "x")  # Scatter plot
rgl.bbox(color = "#333377") # Add bounding box decoration
```

Alternative representation

```{r}
rgl_init()
plot3d(x, y, z, col="blue", type ="p")
```

```{r}
rgl_init()
plot3d(x, y, z, col="blue", box = FALSE,
       type ="s", radius = 0.085)
#ellips <- ellipse3d(cov(cbind(x,y,z)), 
#            centre=c(x, y, z), level = 0.2)
#plot3d(ellips, col = "blue", alpha = 0.2, add = TRUE, box = FALSE)
```

```{r}
rgl.snapshot(filename = "D:/Projects/2022_MANUSCRITOS/2020_SUMMA_BIOL/Notebooks_Programs/Results/Phy_G_3D_Sport.png")
```

```{r}
test_data <- read.csv("D:/Projects/2022_MANUSCRITOS/2020_SUMMA_BIOL/Notebooks_Programs/Results/Phy_S_Diag_Mhealt_x.csv")
plot(test_data$S_M,test_data$Geom_Phi)

```

